sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-12-10 doc/paper.Rmd
a possible way to get von Humboldt in

Ecological data are often collected in space or in space repeatedly over time.
While such data are a rich source of information about ecological processes [@legendre1989a], they are challenging to properly model---data closer in space and time are usually more similar to each other than data farther apart, both due to measured and unmeasured variables [@cressie1993; @cressie2011].
While measured variables can be accounted for via predictors in a model (e.g., measuring and modelling temperature effects on species abundance), unmeasured variables (e.g., everything influencing species abundance but not explicitly modelled) can cause residual spatial correlation.
<!-- ecology has a latitude gradient: von Humboldt, A. (1808) Ansichten der Natur mit wissenschaftlichen Erlauterungen. J.G. Cotta, Tübingen, Germany. -->
Accounting for this residual correlation is important because doing so (1) allows for valid statistical inference [@legendre1989a; @dormann2007], (2) can improve predictions compared to ignoring it [e.g., @shelton2014], and (3) can be of ecological interest itself [e.g., @thorson2019d; @barnett2021].

>>>>

Ecological data are often collected in space or in space repeatedly over time.
While such data are a rich source of information about ecological processes [@legendre1989a], they are challenging to properly model---data closer in space and time are usually more similar to each other than data farther apart, both due to measured and unmeasured variables [@cressie1993; @cressie2011].
While measured variables can be accounted for via predictors in a model (e.g., measuring and modelling temperature effects on species abundance), unmeasured variables (e.g., everything influencing species abundance but not explicitly modelled) can cause residual spatial correlation.
<!-- ecology has a latitude gradient: von Humboldt, A. (1808) Ansichten der Natur mit wissenschaftlichen Erlauterungen. J.G. Cotta, Tübingen, Germany. -->
<!-- While measured variables can be accounted for via predictors in a model (e.g., including latitude, recognized as an important ecological gradient as far back as von Humboldt (1808)), unmeasured variables (e.g., everything influencing species abundance but not explicitly modelled) can still cause residual spatial correlation. -->
Accounting for this residual correlation is important because doing so (1) allows for valid statistical inference [@legendre1989a; @dormann2007], (2) can improve predictions compared to ignoring it [e.g., @shelton2014], and (3) can be of ecological interest itself [e.g., @thorson2019d; @barnett2021].
